Adaboost for face image classification

### 2019-11-14 simple_ml/example/report3/paper.tex
complete report3.

\subsection{Weak learner's performance's effect.}
In this section, we explore the influence of weak learner's performance in Adaboost. We set number of weak learner to five weak learner due to computation limitation. \par 
In this section, we explore the influence of C in SVM. We set $log_{10}^{C}$ in a coarse-grained range e.g. -2, -1, 0, 1, 2, 3, 4, and compare $loss_{val}$,  $accuracy_{val}$ and $l_2norm$ of $w$, where $loss =  \frac{w^Tw}{2} + \frac{C}{m}\Sigma_{i=1}^m \max(0, 1 - y_i(w^Tx_i + b))$ and $ l_2norm(w) = \sqrt{\Sigma_{i=1}^n w_i}$. We zero initilize $w$ and $b$, use SGD as optimizer with learning rate 0.001, set epoch to 200 and batch size to 64. \par
From Fig~\ref{fig:C_loss}, we can see that, when $C$ is very small, e.g. smaller than 1, the gradient descent process is very unstable and the loss is higher than other. As $C$ goes up, the loss decreases, e.g. from 1 to 10. But when $C$ gets larger than 10, the loss become slightly bigger, but still better than small $C$, like 0.01 and 0.1. So, we can draw the conclusion that, bigger $C$ usually better than small $C$, and the best C is in the middle, which is not very big but also not very small. As we can see from Fig~\ref{fig:C_loss}, $C = 10$ gets the lowest loss.   \par
From Fig~\ref{fig:C_acc}, we can see that, when $C$ is very small, e.g. 0.01, the accracy in both training set and validation set is much lower than others. Small $C$ like 0.01 and 0.1 get worse performance comparing with others. In training set, the accracy is very similar when $C$ bigger than 1. In validation set, $C = 10$ gets slightly better performance comparing with others. Best validation performance under each $C$ setting can be found in Table~\ref{tab:C_performance}. \par
From Fig~\ref{fig:c_mag}, we can see that, as $C$ goes up, $l_2norm(w)$ goes up except $C = 100$. As said above, $\frac{1}{2C}$ can be viewed as regurilizer, so when $C$ is bigger, the smaller of its impact on regurization, so the weight maginitude goes up. \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{c_val_loss}
		% Create a subtitle for the figure.
		\caption{$Loss_{val}$ under different $C$.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:C_loss}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{c_val_acc}
		% Create a subtitle for the figure.
		\caption{$Accuracy_{val}$ under different $C$.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:C_acc}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{c_magnitude}
		% Create a subtitle for the figure.
		\caption{The magnitude of $W$ under different $C$.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:c_mag}
	\end{center}
\end{figure} \par

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{Best Result in Validation set under different C's settings. Best result are in bold.}
		\label{tab:C_performance}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|c|c|c|c|c|}
			% To create a horizontal line, type \hline
			\hline
			% To end a column type &
			% For a linebreak type \\
			$log_{10}^{C}$ & -2 & -1 & 0 & 1 & 2 & 3 & 3 \\
			\hline
			$Accuracy_{val}$   & 84.39 & 84.42 & 84.72 & \textbf{84.75} & 84.73 & 84.56 & 84.49   \\
			\hline
			$Epoch$  & \textbf{2} & 86 & 62 & 138 & 173 & 147 & 145   \\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\subsection{Loss function comparision in SVM}
In this section, we conduct experiment on loss function comparision in SVM. Especially, we compare two loss function, hinge loss which is illustrated above And Exponential Loss (EL), where $EL = exp\{-y_i(w^Tx_i + b)\}$. More formaly, $HL(z) = max(0, 1 - z)$ and $EL(z) = exp^{-z}$. Fig~\ref{fig:loss_fn} shows that HL and EL both have one advantage and one disadvantage. HL is stable when z is very mall, however EL become very large, which will lead to numeric unstable. EL is derivable every but HL is not derivable when $z = 1$. We use SGD as optimizer with learning rate 0.001, set epoch to 200, batch size to 64 and C to 10.\par
From Fig~\ref{fig:l_val_loss} and Fig~\ref{fig:l_val_acc}, we can see that the performance of HL is superior. EL has numeric unstablity problem during gradient descent process. \par
\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{loss_fn}
		% Create a subtitle for the figure.
		\caption{The plot of HL and EL.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:loss_fn}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{l_val_loss}
		% Create a subtitle for the figure.
		\caption{$loss_{val}$ under different loss function.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:l_val_loss}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{l_val_acc}
		% Create a subtitle for the figure.
		\caption{$Accuracy_{val}$ under different loss function.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:l_val_acc}
	\end{center}
\end{figure} \par

\subsection{initializer comparision in both LR and SVM}
In this section, we compare the performance of LR and SVM under different initializers. We explore over ones, zeros, xavier uniform and random initializer. Ones initializers makes the initial $w$ to all one and Zeros initializers makes the initial $w$ to all zero. Xavier uniform initializer uniformally initial $w$, but bounded to interval $[-n, n]$, where $n$ is $W$'s dimension. Random initializer initial $w$ from standard normal distribution. We use SGD as optimizer with learning rate 0.001, set epoch to 200, batch size to 64 and C to 10 in SVM. We use HL as SVM's loss function and BCE as LR's loss function.\par
From Fig~\ref{fig:lr_init_val_loss} and Fig~\ref{fig:lr_init_val_acc}, we can see that LR's performance is affected by the initializer. Typically Ones initializer brings the worst performance and Xavier uniform initializer bring the best performance. \par
From Fig~\ref{fig:svm_init_val_loss} and Fig~\ref{fig:svm_init_val_acc}, we can see that SVM is very robust to the initializer, different initializer doesn't disturb the performance. \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{lr_init_val_loss}
		% Create a subtitle for the figure.
		\caption{LR's $BCE_{val}$ under different initializer.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:lr_init_val_loss}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{lr_init_val_acc}
		% Create a subtitle for the figure.
		\caption{LR's $Accuracy_{val}$ under different initializer.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:lr_init_val_acc}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{svm_init_val_loss}
		% Create a subtitle for the figure.
		\caption{SVM's $HL_{val}$ under different initializer.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:svm_init_val_loss}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{svm_init_val_acc}
		% Create a subtitle for the figure.
		\caption{SVM's $Accuracy_{val}$ under different initializer.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:svm_init_val_acc}
	\end{center}
\end{figure} \par

\subsection{Optimizer comparision in both LR and SVM}
In this section, we compare the performance of LR and SVM with different optimizers. We explore over Stochastic Gradient Descent (SGD), Momentum, Adam and RMSProp. We use xavier uniform initializer, set learning rate to 0.001 epoch to 200, batch size to 64 and C to 10 in SVM. We use HL as SVM's loss function and BCE as LR's loss function.\par
From Fig~\ref{fig:lr_optim_val_loss} and Fig~\ref{fig:lr_optim_val_acc}, we can see that LR's performance is affected by the optimizer. Typically SGD converges very slowly; Momentum converges faster than SGD, but not as good as SGD's performance; both Adam and RMSProp converge very fast and get to the best performance.\par
From Fig~\ref{fig:svm_optim_val_loss} and Fig~\ref{fig:svm_optim_val_acc}, we can see that SVM is very robust to the optimizer too, different optimizer didn't disturb the performance except Momentum which converges slower and get worse performance comparing with other optimizers. \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{lr_optim_val_loss}
		% Create a subtitle for the figure.
		\caption{LR's $BCE_{val}$ under different optimizer.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:lr_optim_val_loss}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{lr_optim_val_acc}
		% Create a subtitle for the figure.
		\caption{LR's $Accuracy_{val}$ under different optimizer.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:lr_optim_val_acc}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{svm_optim_val_loss}
		% Create a subtitle for the figure.
		\caption{SVM's $HL_{val}$ under different optimizer.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:svm_optim_val_loss}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{svm_optim_val_acc}
		% Create a subtitle for the figure.
		\caption{SVM's $Accuracy_{val}$ under different optimizer.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:svm_optim_val_acc}
	\end{center}
\end{figure} \par

\subsection{Performance on validation set and test set.}
Now we report the final performance of LR and SVM on test set. For LR, we set optimizer to Adam with learning rate 0.001, batch size to 64, initializer to Xavir uniform intitializer, epoch to 200. For SVM, we set optimizer to Adam with learning rate 0.001, batch size to 64, initializer to Xavir uniform initializer, $C$ to 10, epoch to 200. \par
Although LR is very sensitive to initializer and optimizer, but using the best initializer Xavir uniform initializer and optimizer Adam, it outperform SVM by a little under the same experiment settings and enjoys faster convergence. \par

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{Result on a9a validation set and test set. Best result are in bold.}
		\label{tab:final_performance}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|}
			% To create a horizontal line, type \hline
			\hline
			% To end a column type &
			% For a linebreak type \\
			& LR & SVM \\
			\hline
			$Accuracy_{val}$   & \textbf{84.66} & 84.44   \\
			\hline
			$Accuracy_{test}$   & \textbf{84.98} & 84.88   \\
			\hline
			$Epoch$  & \textbf{37} & 107   \\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\section{Conclusion}
In this report, we explore SVM and logistic regression. Specifically, we explore the hyper-parameter $C$ and loss function comparision in SVM, we draw the conclusion that best $C$ is something in the middle (not too small and too big) and hinge loss is better than exponential loss. Besides that, we also compare different initializers and optimizers in logistic regression. We oberseve that SVM is more robust than logistic regression under differen initializer and optimizer settings; Adam and RMSProp both obtain very good performance in logistic regression and SVM. Finally, we report te performance of LR and SVM under each best settings and get their performance on test set, suprisingly LR outperform SVM in a little performance.\par

>>>>

\subsection{Weak learner's performance's effect.}
In this section, we explore the influence of weak learner's performance in Adaboost. \par 
We first explore the performance of DTC under different depths setting. From Fig~\ref{fig:d_weak_acc} we can see that, as the depth of DTC goes depper, its performance on training set become better but in validation set, the performance of it first gets better and then goes down which indicating the overfitting problem. The best performance in validation set is $89$ occurs when $depth = 4$. \par
Then we explore Adaboost's performance with different weaker learner.  We set number of weak learner to five due to computation limitation. We can see from Fig~\ref{fig:d_acc} that, both Adaboost and DTC suffer from overfitting prolem. But Adaboost outperform DTC under all setting except in validation set when depth equals to 5. We can see that, as the performance of weak learner goes up, the performance of Adaboost goes down, maybe due to the overfitting problem. The intuition is that, if a weak learner's performance is good enough in training set, then Adaboost with more weak learner is prone to overfit the dataset.  \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{d_weak_acc}
		% Create a subtitle for the figure.
		\caption{DTC accuracy under different $depth$.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:d_weak_acc}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{d_acc}
		% Create a subtitle for the figure.
		\caption{Adaboost and DTC accuracy under different $depth$. Note that we set number of weak learner to 5 in Adaboost.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:d_acc}
	\end{center}
\end{figure} \par

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{DTC and Adaboost accuracy on validation set. Best result are in bold.}
		\label{tab:d_dtc_ad}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|c|c|c|}
			\hline
			$depth$  & 1 & 2 & 3 & 4 & 5 \\
			\hline
			$DTC$   & 0.82 & 0.88 & 0.87 & \textbf{0.89} & 0.87   \\
			\hline
			$Adaboost$  & \textbf{0.96} & 0.91 & 0.94 & 0.93 & 0.85 \\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\subsection{Performance}
Now we report the final performance of Adaboost under the best settings e.g. DTC'depth = 3 and number of weak learner = 5. The final result is shown in Fig~\ref{tab:final_performance}.

\begin{table}[!hbt]
	% Center the table
	\begin{center}
		% Title of the table
		\caption{Adaboost's best accuracy. }
		\label{tab:final_performance}
		% Table itself: here we have two columns which are centered and have lines to the left, right and in the middle: |c|c|
		\begin{tabular}{|c|c|c|}
			\hline
			$train$  & $validation$ & $test$  \\
			\hline
			1.00   & 0.98 & 0.97  \\
			\hline
		\end{tabular}
	\end{center}
\end{table} \par

\subsection{Number of weak learner's exploration}
In this section, we conduct experiment on number of weak learner's exploration. \par 
We first set the depth to 1 due to computation limitation. Under this setting, More weak Learner benefits the Adaboost, as shown in Fig~\ref{fig:n_ad_acc}.\par
Then we evalutate the performance of Adaboost under settings that DTC' depth ranges from 1 to 5 and number of weak learner ranges from 1 to 5. As illustrated in Fig~\ref{fig:dn_ad_tr_acc}, in training set, Adaboost benefits from both the depth and the number of weak learners. Note that, as the depth of weak learner goes up, the performance gained by adaboost goes down. As shown in Fig~\ref{fig:dn_ad_val_acc}, when DTC's depth equals to 3, as the number of weak learner goes up, Adaboost has better performance. When DTC's depth equals to 5, Adaboost's performance is the worst. \par
Now, we cann draw the following conclusions: \\
1. Weaker learner gets larger performance gain in Adaboost. \\
2. Very Strong weak learner tends to have overfitting problem in Adaboost. \\
3. More weak learner boost the performance in Adaboost when weak learner is not very strong. \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{n_ad_acc}
		% Create a subtitle for the figure.
		\caption{Adaboost accuracy with different number of DTC ($depth=1$).}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:n_ad_acc}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{dn_ad_tr_acc}
		% Create a subtitle for the figure.
		\caption{Adaboost accuracy with different number of DTC ($depth = 1, 2, 3, 4, 5$) in training set.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:dn_ad_tr_acc}
	\end{center}
\end{figure} \par

\begin{figure}[!hbt]
	% Center the figure.
	\begin{center}
		% Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
		\includegraphics[width=\columnwidth]{dn_ad_val_acc}
		% Create a subtitle for the figure.
		\caption{Adaboost accuracy with different number of DTC ($depth = 1, 2, 3, 4, 5$) in validation set.}
		% Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
		\label{fig:dn_ad_val_acc}
	\end{center}
\end{figure} \par

\section{Conclusion}
In this report, we explore Adaboost algorithm. Specifically, we explore the weak learner's performance and number of weak learner in Adaboost. We draw the conclusions that weaker learner can get more performance gain in Adaboost and stronger learner tends to have overfitting problem in Adaboost. And finally we report the performance on test set under the best validation setting. \par
